---
title: "Accessing the data"
author: "Nino Moriaty"
output: html_notebook
---


# Inspect your data
```{r}
# show the class of your object
class(var)
# show the dimension of a variable, especially for data frame and data table.
dim(df)
```


# Data Types / Mode
## Numeric data (Integer, Float and Double and Complex (imaginary numbers) )
## Character data / String
## Logical data
## Raw (bytes).



# Data Structure / Class
## Vector (1D)
Tips: One vector contains one data type/mode, including numeric data, character data, or logical data
Tips: Scalars are one-element vectors. However, they are mostly settled as constrants in R language.
```{r}
# Create a vector
vector1 <- c(1, 2, 3)
vector2 <- c(0, 1, 2)
vector3 <- c(0, 0, 1)

# transform


# Refer to elements of a vector using a numeric vector
vector1[1] # Tips: start from 1 rather than 0
vector1[1:2]
vector1[c(1:2)]
vector1[c(1,2,3)] 


```

## Factors (Special)
Tips: factors store nominal or ordinal variables

## Matrix (2D)
```{r}
# Create Matrices
matrix1 <- matrix(vector1, vector2)
matrix1 <- matrix(1:20, nrow=5, ncol=4)

# transform the data sturcture
as.matrix() # plus a as. before the type of data structrue
matrix()


```

## Array (Multi-dimensions)
```{r}
# Create Arrays
array1 <- array(vector1, vector2, vector3)


```

## Data Frame (2.5D)
Tips: `df` can store all kinds of *data types*
```{r}
# Create data frame
df <- data.frame()
df <- data.frame(matrix1, stringsAsFactors = FALSE) # stringAsFactors sometimes cause errors when reading strings in the df 

# transformation
df <- as.data.frame(x)

# Create an emplty data frame with colnames or rownames
data.frame(matrix(nrow=2,ncol=2))

# assign rownames and colnames to df
rownames(df) <- c("")
colnames(df) <- c("")
```

## Data Table
Tips: `dt` is a special data frame for `dplyr`
```{r}
data.table()
```

## List
Tips: `List` can contain all kins of *data structures*
```{r}
# Create a list
ls1 <- list(vector1, matrix1)

# Extract factors in the list
ls1[[1]]
ls1

#


```

## Define your own class
Tips: More information could be seen in the R Package Section in the Wiznote.


# Input data
## Files
```{r}
# Refer to other scripts
source("pahtname") # if the file is in the current workspace, the filename could be used directly.

# Input data
table <- read.table(file) # data


# data from package
iris <- data("Arthritis") # from pkg "vcd"


```

## Website




# Annotating datasets

















